---
title: "Chapter 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# 5 Data transformation
## 5.1 Introduction
### 5.1.1 Prerequisites

```{r}
library(nycflights13)
library(tidyverse)
```

### 5.1.2 nycflights13
```{r}
flights
```

### 5.1.3 dplyr basics
## 5.2 Filter rows with filter()

```{r}
filter(flights, month == 1, day == 1)
```

```{r}
jan1 <- filter(flights, month == 1, day == 1)
```

```{r}
(dec25 <- filter(flights, month == 12, day == 25))
```

### 5.2.1 Comparisons
```{r, eval = F}
filter(flights, month = 1)
```

```{r}
sqrt(2) ^ 2 == 2
1 / 49 * 49 == 1
```

```{r}
near(sqrt(2) ^ 2,  2)
near(1 / 49 * 49, 1)
```

### 5.2.2 Logical operators

```{r}
filter(flights, month == 11 | month == 12)
```

```{r}
nov_dec <- filter(flights, month %in% c(11, 12))
```

```{r}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

### 5.2.3 Missing values

```{r}
NA > 5
#> [1] NA
10 == NA
#> [1] NA
NA + 10
#> [1] NA
NA / 2
#> [1] NA
```

```{r}
NA == NA
#> [1] NA
```

```{r}
# Let x be Mary's age. We don't know how old she is.
x <- NA

# Let y be John's age. We don't know how old he is.
y <- NA

# Are John and Mary the same age?
x == y
#> [1] NA
# We don't know!
```

```{r}
is.na(x)
#> [1] TRUE
```

```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
#> # A tibble: 1 x 1
#>       x
#>   <dbl>
#> 1     3
filter(df, is.na(x) | x > 1)
#> # A tibble: 2 x 1
#>       x
#>   <dbl>
#> 1    NA
#> 2     3
```

### 5.2.4 Exercises

1.1 Had an arrival delay of two or more hours

```{r}
flights %>% filter(arr_delay >= 120)
```

1.4 Departed in summer (July, August, and September)

```{r}
flights %>% filter(month %in% c(7,8,9))
```

1.5 Arrived more than two hours late, but didn’t leave late

```{r}
flights %>%
  filter(arr_delay > 120) %>%
  filter(dep_delay <= 0)
```

1.7 Departed between midnight and 6am (inclusive)

```{r}
flights %>% filter(dep_time >= 0 & dep_time <= 600)
```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
flights %>% filter(between(dep_time, 0, 600))
```
3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
flights %>% filter(is.na(dep_time))
```
Canceled flights

### 5.3.1 Exercises

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
flights %>% arrange(!is.na(.))
```


2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
flights %>% arrange(desc(arr_delay))
flights %>% arrange(dep_delay)
```

3. Sort flights to find the fastest (highest speed) flights.
```{r}
flights %>% arrange(desc(distance/air_time))
flights %>% mutate(speed = distance/air_time) %>% arrange(desc(speed))
```

4. Which flights travelled the farthest? Which travelled the shortest?
```{r}
flights %>% arrange(desc(distance))
flights %>% arrange(distance)
```

### 5.4.1 Problems 1-4

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.


2. What happens if you include the name of a variable multiple times in a select() call?

```{r}
flights %>% select(day, month, day)
```


3. What does the any_of() function do? Why might it be helpful in conjunction with this vector?

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
```

```{r}
flights %>% select(any_of(vars))
```


4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

```{r}
select(flights, contains("TIME", ignore.case = F))
```

### 5.5.2 Problems 1-5
1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
flights %>% transmute(dep_time,
                      dep_mins = (dep_time %/% 100 * 60) + dep_time %% 100,
                      sched_dep_time,
                      sched_dep_mins = (sched_dep_time %/% 100 * 60) + sched_dep_time %% 100)
```

2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
flights %>% transmute(
  air_time,
  arr_time,
  dep_time,
  dep_mins = (dep_time %/% 100 * 60) + dep_time %% 100,
  arr_mins = (arr_time %/% 100 * 60) + arr_time %% 100,
  diff_mins = arr_mins - dep_mins
)
```

```{r}
flights %>% transmute(
  air_time,
  arr_time,
  dep_time,
  diff_time = arr_time - dep_time,
  diff_mins = (diff_time %/% 100 * 60) + diff_time %% 100,
  areSame = air_time == diff_mins
)
```



3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

dep_delay = dep_time - sched_dep_time

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
flights %>% filter(rank(dplyr::desc(arr_delay), ties.method = "min")<=10) 
```

```{r}
flights %>% filter(rank(dplyr::desc(arr_delay), ties.method = "first")<=10)  %>%
  arrange(desc(arr_delay))

```


5. What does 1:3 + 1:10 return? Why?

```{r}
1:3 + 1:10
```

### 5.6.7 Problem 5
5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n())

```{r}
flights %>% group_by(carrier, dest) %>% summarise(n())
```


```{r}
flights %>%
  group_by(carrier) %>%
  summarize(avg_delay = mean(arr_delay, na.rm = T)) %>%
  arrange(desc(avg_delay)) %>%
  left_join(airlines) %>%
  select(name, everything())
```

```{r}
flights %>%
  group_by(carrier, dest) %>%
  summarize(avg_delay = mean(arr_delay, na.rm = T)) %>%
  arrange(desc(avg_delay)) %>%
  left_join(airlines) %>%
  select(name, everything())
```

```{r}
flights %>%
  group_by(carrier, origin, dest) %>%
  summarize(avg_delay = mean(arr_delay, na.rm = T)) %>%
  arrange(desc(avg_delay)) %>%
  left_join(airlines) %>%
  select(name, everything())
```

```{r}
flights %>%
  group_by(carrier, origin, dest) %>%
  summarize(avg_delay = mean(arr_delay, na.rm = T)) %>%
  arrange(desc(avg_delay)) %>%
  left_join(airlines) %>%
  ungroup() %>%
  group_by(name) %>%
  summarize(avg_delay = mean(avg_delay, na.rm = T)) %>%
  arrange(desc(avg_delay))
```

