---
title: "Chapter 7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

### 7.3.4 Exercises

1. Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
head(diamonds)
```

```{r}
diamonds %>%
  pivot_longer(x:z, names_to = "dimension", values_to = "mm") %>%
  ggplot(aes(x = mm, fill = dimension)) +
  geom_histogram() +
  facet_wrap(~dimension, scales = "free_x")
  
```


2. Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
diamonds %>%
  ggplot(aes(x=price)) +
  geom_histogram(binwidth = 1000)
```

```{r}
diamonds %>%
  mutate(carat = round(carat)) %>%
  ggplot(aes(x=price)) +
  geom_histogram() +
  facet_wrap(~carat, scales = "free_x")
```


3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r}
sum(diamonds$carat==0.99)
sum(diamonds$carat==1)
```


4. Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?

```{r}
diamonds %>%
  ggplot(aes(x=carat)) +
  geom_histogram(binwidth = 0.2)
```

```{r}
diamonds %>%
  ggplot(aes(x=carat)) +
  geom_histogram(binwidth = 0.2) +
  xlim(0,3)
```

```{r}
diamonds %>%
  ggplot(aes(x=carat)) +
  geom_histogram(binwidth = 0.2) +
  coord_cartesian(xlim = c(0,3))
```

### 7.4.1 Exercises

1. What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?

```{r}
diamonds2 <- diamonds %>% 
  mutate(carat = ifelse(carat < 3, round(carat), NA))
```

```{r}
diamonds2 %>%
  ggplot(aes(x = carat)) +
  geom_histogram()
```

```{r}
diamonds2 %>%
  ggplot(aes(x = carat)) +
  geom_bar()
```


```{r}
diamonds2 %>%
  ggplot(aes(x = as.factor(carat))) +
  geom_bar()
```

2. What does na.rm = TRUE do in mean() and sum()?

```{r}
mean(diamonds2$carat)
mean(diamonds2$carat, na.rm = T)
```

```{r}
sum(diamonds2$carat)
sum(diamonds2$carat, na.rm = T)
```

### 7.5.1.1 Exercises

1. Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights.

```{r}
library(nycflights13)
flights %>%
  mutate(cancelled = is.na(dep_time)) %>%
  ggplot(aes(x = sched_dep_time, color = cancelled)) +
  geom_freqpoly()
```


2. What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?

```{r}
colnames(diamonds)
cor(diamonds[,-c(2,3,4,7)], diamonds$price)
```

```{r}
diamonds %>%
  ggplot(aes(x=cut, y = price)) +
  geom_boxplot()
```

```{r}
diamonds %>%
  ggplot(aes(x=color, y = price)) +
  geom_boxplot()
```


```{r}
diamonds %>%
  ggplot(aes(x=clarity, y = price)) +
  geom_boxplot()
```

```{r}
diamonds %>%
  ggplot(aes(x=carat, y = price, color = cut)) +
  geom_point()
```


3. Install the ggstance package, and create a horizontal boxplot. How does this compare to using coord_flip()?

```{r}
library(ggstance)
ggplot(diamonds, aes(x = cut, y = carat)) + geom_boxplot() + coord_flip()
```

```{r}
ggplot(diamonds) + geom_boxploth(aes(x = carat, y = cut))
```

5. Compare and contrast geom_violin() with a facetted geom_histogram(), or a coloured geom_freqpoly(). What are the pros and cons of each method?

```{r}
ggplot(diamonds,aes(x = cut, y = carat)) + geom_violin()
ggplot(diamonds,aes(x = carat)) + geom_histogram() + facet_wrap(~cut, nrow = 1)
ggplot(diamonds,aes(x = carat, color = cut)) + geom_freqpoly()
```


6. If you have a small dataset, it’s sometimes useful to use geom_jitter() to see the relationship between a continuous and categorical variable. The ggbeeswarm package provides a number of methods similar to geom_jitter(). List them and briefly describe what each one does.